MSD panels 1, 2 and 3 Explant Sup

### 2016-08-23 MSD_panels_1_2_3_HVE_sup_analysis.Rmd
Changed the layout of the individual plots

library(dplyr)
library(ggplot2)
library(pander)
library(reshape2)
library(stringr)

@@

HVE_all_panels <- HVE_all_panels %>%
  select(Assay,Sample, Sample.Group, Plate.Name, X..Recovery, Concentration, Calc..Concentration, Detection.Range, Fit.Statistic..RSquared,Spot)
```

@@

The following plots show that the 1:2 dilution of the samples resulted in more samples being in "In Detection Range." Going forward I'll only show data from the 1:2 dilutions.

@@

HVE_all_panels_noStds_Below <- HVE_all_panels_noStds%>%
  mutate(Panel = ifelse(str_detect(Plate.Name, "panel_1"),"Panel 1",ifelse(str_detect(Plate.Name,"panel_2"),"Panel 2","Panel 3")))%>%
 group_by(Plate.Name,Sample_Dilution,Panel,Assay, Detection.Range)%>%
  summarise(Number_of_Samples = n())

@@

ggplot(HVE_all_panels_noStds_Below, aes(x = Assay, y=Number_of_Samples))+
         geom_point(aes(color = Detection.Range),size = 2)+
  theme(axis.text.x = element_text(angle=25, size = 8))+
  labs(y = "Number of Samples", x = "Analyte")+
  facet_wrap(Sample_Dilution~Panel, scales = "free")+
  ggtitle("Number of Samples in each Detection Range per Analyte")
```

@@

### NOTE ### I'm going to filter for just the 1:2 dilution samples here since they gave more in-detection.

HVE_all_panels_noStds_1to2<-HVE_all_panels_noStds %>%
  filter(Sample_Dilution =="1:2")

HVE_all_panels_noStds_1to2_noBelow <- HVE_all_panels_noStds_1to2 %>%
  filter(Detection.Range != "Below Fit Curve Range" & Detection.Range != "Below Detection Range")

#averaging the replicates
avgConc_HVE_all_panels_noStds_1to2_noBelow <- HVE_all_panels_noStds_1to2_noBelow %>%
group_by(Sample, Assay, Donor, Timepoint,Virus)%>%
summarize(avgConc = mean(Calc..Concentration))%>%
  ungroup()

@@

## Plots of analyte concentrations

Black points show the means.

@@

Also note:
Some of the samples had one rep that was below detection and another that was in range. For these samples I used the remaining "in range" rep as the "average" for that sample.

```{r  samples plots, echo=FALSE, message = FALSE, warning = FALSE}

@@

analyteList<-as.list(unique(avgConc_HVE_all_panels_noStds_1to2_noBelow_fullSet$Assay))

#write a function to make a plot for each analyte

plotAnalytes <- function (Analyte){
   avgConc_HVE_all_panels_noStds_1to2_noBelow_fullSet %>%
     filter(Assay == Analyte)%>%

     ggplot( aes(x = Timepoint, y = avgConc))+

   geom_point(aes(color = Donor ), alpha = 0.5, size = 3)+
  geom_line(aes(group = Donor, color = Donor), alpha = 0.5)+
    stat_summary(fun.y = mean, geom = "point")+
    stat_summary(fun.y = mean, geom = "line",aes(group = 1))+

   scale_shape_manual(values = c("Above Fit Curve Range" = 1, "Below Detection Range" = 1, "In Detection Range" = 16))+

   scale_y_log10()+

   labs( y = "Concentration")+

  facet_wrap(~Virus, scales = "free_x")+

   theme(axis.text.x = element_text(angle = 25, size = 9))+
     ggtitle(Analyte)
}

lapply(analyteList, FUN = plotAnalytes)

>>>>

library(dplyr)
library(ggplot2)
library(pander)
library(reshape2)
library(stringr)
library(swarm)

@@

HVE_all_panels <- HVE_all_panels %>%
  select(Assay,Sample, Sample.Group, Plate.Name, X..Recovery, Concentration, Calc..Concentration, Detection.Range, Fit.Statistic..RSquared,Spot)

@@

The following plots show that the 1:2 dilution of the samples resulted in more samples being in "In Detection Range."

@@

HVE_all_panels_noStds_NumInEach <- HVE_all_panels_noStds%>%
  mutate(Panel = ifelse(str_detect(Plate.Name, "panel_1"),"Panel 1",ifelse(str_detect(Plate.Name,"panel_2"),"Panel 2","Panel 3")))%>%
 group_by(Plate.Name,Sample_Dilution,Panel,Assay, Detection.Range)%>%
  summarise(Number_of_Samples = n())

@@

ggplot(HVE_all_panels_noStds_NumInEach, aes(x = Assay, y=Number_of_Samples))+
         geom_point(aes(color = Detection.Range),size = 2)+
  theme(axis.text.x = element_text(angle=25, size = 8))+
  labs(y = "Number of Samples", x = "Analyte")+
  facet_wrap(Sample_Dilution~Panel, scales = "free")+
  ggtitle("Number of Samples in each Detection Range per Analyte")
```

```{r multiplying by dilution factor, echo=FALSE, message = FALSE, warning = FALSE}

#I'm going to make a column for estimated stock concentration by multiplying the avgConc by 2 or 5 (since I did a 1:2 and 1:5 dilution on the samples) so I can better compare with the explant sup concentrations.

HVE_all_panels_noStds <- HVE_all_panels_noStds%>%
  mutate(Estimated_Stock_Conc = ifelse(
    Sample_Dilution == "1:2", Calc..Concentration * 2,
    ifelse(Sample_Dilution == "1:5", Calc..Concentration *5, NA)
  ))

```{r exclude belows, echo=FALSE, message = FALSE, warning = FALSE}

HVE_all_panels_noStds_noBelow <- HVE_all_panels_noStds %>%
  filter(Detection.Range != "Below Fit Curve Range" & Detection.Range != "Below Detection Range")
```

## All Samples Estimated Stock Concentrations

* I multiplied by the dilution factors so this is an estimated concentration of the stock samples.

* Note that this is on a log(10) scale; most concentrations are a lot lower than the high ones.

```{r  all sample concentrations, echo=FALSE, message = FALSE, warning = FALSE}

#Plot of all samples and their concentrations

ggplot(HVE_all_panels_noStds_noBelow,aes( x = Timepoint, y = Estimated_Stock_Conc))+
         geom_point(aes(color = Assay),position = position_swarm (0.5),size = 2)+
  scale_y_log10()+
  labs(y = "Estimated Stock Concentration pg/ml")+
  facet_wrap(~Virus)

```{r filter for 1:2, echo=FALSE, message = FALSE, warning = FALSE}
### NOTE ### I'm going to filter for just the 1:2 dilution samples here since they gave more in-detection.

HVE_all_panels_noStds_1to2_noBelow<-HVE_all_panels_noStds_noBelow %>%
  filter(Sample_Dilution =="1:2")

@@

#averaging the replicates
avgConc_HVE_all_panels_noStds_1to2_noBelow <- HVE_all_panels_noStds_1to2_noBelow %>%
group_by(Sample, Assay, Donor, Timepoint,Virus)%>%
summarize(avgEstConc = mean(Estimated_Stock_Conc))%>%
  ungroup()

@@

## Plots of analyte concentrations
Since the 1:2 diluted samples were more likely to be in the detection range, I am only showing data for those.

@@

Also note:
If a sample had one rep that was below detection and another that was in range, I used the remaining "in range" rep as the "average" for that sample.

```{r  individual analyte plots , echo=FALSE, message = FALSE, warning = FALSE}

ggplot(avgConc_HVE_all_panels_noStds_1to2_noBelow_fullSet, aes(x = Timepoint, y = avgEstConc))+
   geom_point(aes(color = Donor), alpha = 0.75, size = 2)+
  scale_y_log10()+
   labs( y = "Estimated Stock Concentration pg/ml")+
   facet_grid(Assay~Virus, scales = "free_x")+
   theme(axis.text.x = element_text(size = 11))

################## another way of making the plots #########

@@

# analyteList<-as.list(unique(avgConc_HVE_all_panels_noStds_1to2_noBelow_fullSet$Assay))

# #write a function to make a plot for each analyte
# 
# plotAnalytes <- function (Analyte){
#    avgConc_HVE_all_panels_noStds_1to2_noBelow_fullSet %>%
#      filter(Assay == Analyte)%>%
#      
#      ggplot( aes(x = Timepoint, y = avgEstConc))+
#    geom_point(aes(color = Donor ), alpha = 0.5, size = 3)+
#   geom_line(aes(group = Donor, color = Donor), alpha = 0.5)+
#     stat_summary(fun.y = mean, geom = "point")+
#     stat_summary(fun.y = mean, geom = "line",aes(group = 1))+
#    scale_shape_manual(values = c("Above Fit Curve Range" = 1, "Below Detection Range" = 1, "In Detection Range" = 16))+
#    scale_y_log10()+
#    labs( y = "Concentration pg/ml")+
#   facet_wrap(~Virus, scales = "free_x")+
#    theme(axis.text.x = element_text(size = 11))+
#      ggtitle(Analyte)
# }
# 
# 
# 
# lapply(analyteList, FUN = plotAnalytes)

### 2016-08-23 MSD_panels_1_2_3_explant_sup_analysis.Rmd
Changed the layout of the individual plots

#looking at which samples have STDs below the fit curve or below the detection range (i.e. you can't get numbers for % recovery) 
explant_stds_all_panels_below<- explant_stds_all_panels%>%
  filter(Detection.Range == "Below Fit Curve Range" | Detection.Range == "Below Detection Range") %>%
  group_by(Sample)%>%
  summarise("Number of Samples Below Fit Curve Range/Detection Range" = n())

@@

#Split samples names on the "-"
donorTimeVirus<-colsplit(explant_all_panels_noStds$Sample,"-", c("Donor","Timepoint","Virus") )

#bind the split col back to allPanels

explant_all_panels_noStds<-cbind(donorTimeVirus, explant_all_panels_noStds)

@@

#What number of samples are in each category of detection?
explant_all_panels_noStds_Below <- explant_all_panels_noStds%>%
  mutate(Panel = ifelse(str_detect(Plate.Name, "panel_1"),"Panel 1",ifelse(str_detect(Plate.Name,"panel_2"),"Panel 2","Panel 3")))%>%
 group_by(Panel,Assay, Detection.Range)%>%
  summarise(Number_of_Samples = n())

@@

ggplot(explant_all_panels_noStds_Below, aes(x = Assay, y=Number_of_Samples))+
         geom_point(aes(color = Detection.Range),size = 2)+
  theme(axis.text.x = element_text(angle=25, size = 8))+
  labs(y = "Number of Samples", x = "Analyte")+
  facet_wrap(~Panel, scales = "free")+
  ggtitle("Number of Samples in each Detection Range per Analyte")

@@

```{r avg reps, echo=FALSE, message = FALSE, warning = FALSE}

@@

explant_all_panels_noStds_noBelow <- explant_all_panels_noStds %>%
  filter(Detection.Range != "Below Fit Curve Range" & Detection.Range != "Below Detection Range")

#averaging the replicates
avgConc_explant_all_panels_noStds_noBelow <- explant_all_panels_noStds_noBelow %>%
group_by(Sample, Assay, Donor, Timepoint,Virus)%>%
summarize(avgConc = mean(Calc..Concentration))%>%
  ungroup()

@@

Also note:
Some of the samples had one rep that was below detection and another that was in range. For these samples I used the remaining "in range" rep as the "average" for that sample.

```{r  samples plots, echo=FALSE, message = FALSE, warning = FALSE}

#make a list of the analytes

#I'm using a unique vector here instead of the levels of Assay because I don't want to include IFNa2a OR Il-7. There weren't any data points in the detection range for those analytes (so I don't want to try to make a plot of it)

analyteList<-as.list(unique(avgConc_explant_all_panels_noStds_noBelow$Assay))

#write a function to make a plot for each analyte

plotAnalytes <- function (Analyte){
   avgConc_explant_all_panels_noStds_noBelow_fullSet %>%
     filter(Assay == Analyte)%>%

     ggplot( aes(x = Timepoint, y = avgConc))+

   geom_point(aes(color = Donor ), alpha = 0.5, size = 3)+
  geom_line(aes(group = Donor, color = Donor), alpha = 0.5)+
    stat_summary(fun.y = mean, geom = "point")+
    stat_summary(fun.y = mean, geom = "line",aes(group = 1))+

   scale_shape_manual(values = c("Above Fit Curve Range" = 1, "Below Detection Range" = 1, "In Detection Range" = 16))+

   scale_y_log10()+

   labs( y = "Concentration")+

  facet_wrap(~Virus, scales = "free_x")+

   theme(axis.text.x = element_text(angle = 25, size = 9))+
     ggtitle(Analyte)
}

lapply(analyteList, FUN = plotAnalytes)

>>>>

#looking at which samples have STDs below the fit curve or below the detection range (i.e. you can't get numbers for % recovery)

explant_stds_all_panels_below<- explant_stds_all_panels%>%
  filter(Detection.Range == "Below Fit Curve Range" | Detection.Range == "Below Detection Range") %>%
  group_by(Sample)%>%
  summarise("Number of Samples Below Fit Curve Range/Detection Range" = n())

@@

#Split samples names on the "-"
donorTimeVirus<-colsplit(explant_all_panels_noStds$Sample,"-", c("Donor","Timepoint","Virus"))

#Add a column for Sample dilution (all samples were diluted 1:2)

donorTimeVirusDil <- donorTimeVirus %>%
  mutate(Sample_Dilution = "1:2")

#bind the donorTimeVirusDil to allPanels

explant_all_panels_noStds<-cbind(donorTimeVirusDil, explant_all_panels_noStds)

@@

#What number of samples are in each category of detection?
explant_all_panels_noStds_NumInEach <- explant_all_panels_noStds%>%
  mutate(Panel = ifelse(str_detect(Plate.Name, "panel_1"),"Panel 1",ifelse(str_detect(Plate.Name,"panel_2"),"Panel 2","Panel 3")))%>%
 group_by(Panel,Assay, Detection.Range)%>%
  summarise(Number_of_Samples = n())

@@

ggplot(explant_all_panels_noStds_NumInEach, aes(x = Assay, y=Number_of_Samples))+
         geom_point(aes(color = Detection.Range),size = 2)+
  theme(axis.text.x = element_text(angle=25, size = 7))+
  labs(y = "Number of Samples", x = "Analyte")+
  facet_wrap(~Panel, scales = "free")+
  ggtitle("Number of Samples in each Detection Range per Analyte")

@@

```{r multiplying by dilution factor, echo=FALSE, message = FALSE, warning = FALSE}

#I'm going to make a column for estimated stock concentration by multiplying the avgConc by 2 (since I did a 1:2 dilution on the samples) so I can better compare with the HVE sup concentrations.

explant_all_panels_noStds <- explant_all_panels_noStds%>%
  mutate(Estimated_Stock_Conc =  Calc..Concentration * 2)

```{r exclude belows, echo=FALSE, message = FALSE, warning = FALSE}

explant_all_panels_noStds_noBelow <- explant_all_panels_noStds %>%
  filter(Detection.Range != "Below Fit Curve Range" & Detection.Range != "Below Detection Range")
```

## All Samples Estimated Stock Concentrations

* I multiplied by the dilution factor so this is an estimated concentration of the stock samples.

* Note that this is on a log(10) scale; most concentrations are a lot lower than the high ones.

```{r  all sample concentrations, echo=FALSE, message = FALSE, warning = FALSE}

#Plot of all samples and their concentrations

ggplot(explant_all_panels_noStds_noBelow,aes( x = Timepoint, y = Estimated_Stock_Conc))+
         geom_point(aes(color = Assay),position = position_swarm (0.5),size = 2)+
  scale_y_log10()+
  labs(y = "Estimated Stock Concentration pg/ml")+
  facet_wrap(~Virus)

```{r averaging reps, echo=FALSE, message = FALSE, warning = FALSE}

#averaging the replicates

@@

avgConc_explant_all_panels_noStds_noBelow <- explant_all_panels_noStds_noBelow %>%
group_by(Sample, Assay, Donor, Timepoint,Virus)%>%
summarize(avgEstConc = mean(Estimated_Stock_Conc))%>%
  ungroup()

@@

Also note:
If a sample had one rep that was below detection and another that was in range, I used the remaining "in range" rep as the "average" for that sample.

```{r  individual analyte plots, echo=FALSE, message = FALSE, warning = FALSE}

ggplot(avgConc_explant_all_panels_noStds_noBelow_fullSet, aes(x = Timepoint, y = avgEstConc))+
   geom_point(aes(color = Donor),position = position_swarm(0.07), alpha = 0.75, size = 1.5)+
  scale_y_log10()+
   labs( y = "Estimated Stock Concentration pg/ml")+
   facet_grid(Assay~Virus, scales = "free_x")+
   theme(axis.text.x = element_text(size = 11))

################## another way of making the plots #####

# #make a list of the analytes
# 
# #I'm using a unique vector here instead of the levels of Assay because I don't want to include IFNa2a OR Il-7. There weren't any data points in the detection range for those analytes (so I don't want to try to make a plot of it)
# 
# 
# analyteList<-as.list(unique(avgConc_explant_all_panels_noStds_noBelow$Assay))
# 
# 
# 
# 
# #write a function to make a plot for each analyte
# 
# plotAnalytes <- function (Analyte){
#    avgConc_explant_all_panels_noStds_noBelow_fullSet %>%
#      filter(Assay == Analyte)%>%
#      
#      ggplot( aes(x = Timepoint, y = Estimated_Stock_Conc))+
#    
#    geom_point(aes(color = Donor ), alpha = 0.5, size = 3)+
#   geom_line(aes(group = Donor, color = Donor), alpha = 0.5)+
#     stat_summary(fun.y = mean, geom = "point")+
#     stat_summary(fun.y = mean, geom = "line",aes(group = 1))+
#    
#    scale_shape_manual(values = c("Above Fit Curve Range" = 1, "Below Detection Range" = 1, "In Detection Range" = 16))+
#   
#    scale_y_log10()+
#    
#    labs( y = "Concentration pg/ml")+
#    
#   facet_wrap(~Virus, scales = "free_x")+
#    
#    theme(axis.text.x = element_text(size = 11))+
#      ggtitle(Analyte)
# }
# 
# 
# 
# lapply(analyteList, FUN = plotAnalytes)
